Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit

### 2024-06-24 optim.qmd
update appendix
add definition of porjecion distance

@@

```{=tex}
\textcolor{orange}{See whether further details are needed for the nonparametric approach.}
```
TODO: add equation reference to main text line 705

@@

## Speed

The speed of optimizing an index function can be measured empirically by
recording the duration of the optimisation. Alternatively, one can also
gauge the speed in terms of computational complexity of evaluating the
index function (for instance, in big O notation, with respect to the
sample size) of computing the index function.

@@

@tbl-smoothness-squintability presents the parameters estimated from the
Gaussian process (variance, range, smooth, and nugget) and the scaled
logistic function ($\theta_1$ to $\theta_4$) for each case. The column
"smooth" is used as the smoothness measure and the column "squint" is
calculated as \[TODO: equation reference\] as squintability. The low
squint value of MIC and TIC index is due to its convex shape of the
index value against the projection distance, as shown in
@fig-idx-proj-dist. Comparing to other concave shapes, the maximum first
derivative happens at a smaller projection distance ($\theta_2$),
requiring the optimiser to get closer to the optimal to see a
significant change in the index value, hence a small squintability
measure.

>>>>

Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$ the optimal basis achieving the maximum index value for a given projection pursuit problem defined by the shape to find, data dimension, and the index function $f(.)$. The projection distance $d(A, A^*)$, is defined as the Frobenius norm $\lVert . \rVert _F$, between the squared difference of these two bases:

$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}$$ {#eq-proj-dist}

The index value $f(XA)$ is defined as described in the introduction.

@@

\textcolor{orange}{See whether further details are needed for the nonparametric approach.}

@@

<!-- ## Speed -->

<!-- The speed of optimizing an index function can be measured empirically by -->
<!-- recording the duration of the optimisation. Alternatively, one can also -->
<!-- gauge the speed in terms of computational complexity of evaluating the -->
<!-- index function (for instance, in big O notation, with respect to the -->
<!-- sample size) of computing the index function. -->

@@

@tbl-smoothness-squintability presents the parameters estimated from the
Gaussian process (variance, range, smooth, and nugget) and the scaled
logistic function ($\theta_1$ to $\theta_4$) for each case. The column
"smooth" is used as the smoothness measure and the column "squint" is
calculated as equation
(@eq-squintability-parametric) as squintability. The low
squint value of MIC and TIC index is due to its convex shape of the
index value against the projection distance, as shown in
@fig-idx-proj-dist. Comparing to other concave shapes, the maximum first
derivative happens at a smaller projection distance ($\theta_2$),
requiring the optimiser to get closer to the optimal to see a
significant change in the index value, hence a small squintability
measure.

### 2024-06-24 optim.tex
add definition of porjecion distance

\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, opacityback=0, bottomrule=.15mm, toprule=.15mm, bottomtitle=1mm, left=2mm, rightrule=.15mm, coltitle=black, titlerule=0mm, toptitle=1mm, opacitybacktitle=0.6, leftrule=.75mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, colback=white]

@@

@@

TODO: add equation reference to main text line 705

@@

\subsection{Speed}\label{speed}

The speed of optimizing an index function can be measured empirically by
recording the duration of the optimisation. Alternatively, one can also
gauge the speed in terms of computational complexity of evaluating the
index function (for instance, in big O notation, with respect to the
sample size) of computing the index function.

@@

Table~\ref{tbl-smoothness-squintability} presents the parameters
estimated from the Gaussian process (variance, range, smooth, and
nugget) and the scaled logistic function (\(\theta_1\) to \(\theta_4\))
for each case. The column ``smooth'' is used as the smoothness measure
and the column ``squint'' is calculated as {[}TODO: equation
reference{]} as squintability. The low squint value of MIC and TIC index
is due to its convex shape of the index value against the projection
distance, as shown in Figure~\ref{fig-idx-proj-dist}. Comparing to other
concave shapes, the maximum first derivative happens at a smaller
projection distance (\(\theta_2\)), requiring the optimiser to get
closer to the optimal to see a significant change in the index value,
hence a small squintability measure.

>>>>

\begin{tcolorbox}[enhanced jigsaw, leftrule=.75mm, rightrule=.15mm, bottomrule=.15mm, colback=white, bottomtitle=1mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, left=2mm, breakable, coltitle=black, toptitle=1mm, titlerule=0mm, colframe=quarto-callout-note-color-frame, title={Algorithm: Jellyfish Optimizer Pseudo Code}]

@@

Let \(A \in \mathbf{R}^{p \times 2}\) be the 2D projection basis and
\(A^*\) the optimal basis achieving the maximum index value for a given
projection pursuit problem defined by the shape to find, data dimension,
and the index function \(f(.)\). The projection distance \(d(A, A^*)\),
is defined as the Frobenius norm \(\lVert . \rVert _F\), between the
squared difference of these two bases:

\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}}\end{equation}

The index value \(f(XA)\) is defined as described in the introduction.

@@

@@

@@

Table~\ref{tbl-smoothness-squintability} presents the parameters
estimated from the Gaussian process (variance, range, smooth, and
nugget) and the scaled logistic function (\(\theta_1\) to \(\theta_4\))
for each case. The column ``smooth'' is used as the smoothness measure
and the column ``squint'' is calculated as equation
(\ref{eq-squintability-parametric}) as squintability. The low squint
value of MIC and TIC index is due to its convex shape of the index value
against the projection distance, as shown in
Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
maximum first derivative happens at a smaller projection distance
(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
see a significant change in the index value, hence a small squintability
measure.
